---
title: "739 Project"
author: "Hugh O'Hara, Megan Parmer, Devin Rohler"
date: "11/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load libraries
library(tidyverse)
library(janitor)
library(lubridate)
library(GGally)
library(psych)
library(corrplot)
library(caTools)
library(forecast)
library(easystats)
library(scales)
```

```{r}
# import
setwd("/Users/Devin/Desktop/School Work/Graduate/Fall 2022/Advanced Datamining")
car= read.csv("car_purchasing.csv", encoding = "UTF-8")
```

***Project Explanation***
This project will be using a dataset with information on vehicle sales from a wide variety of people and attirbutes.  This data set can bring in lots of information about marketing, sales and much more. A lot of insight can be gained for car salespersons, car manufacturers, loan agencies and marketing specialists.
We will be performing machine learning models to see how accuretely we can predict car purchase price based on a customer's attributes. In terms of car salespersons, they could use this model to target certain customers.  Marketing and sales tactics can benefit very much from being able to know exactly what type of customer to go after for certain valued cars. This will save time, effort and money. There is a lot more insight that can be gained once worked through exploration and modeling. 


***Data Cleansing***

```{r}
# clean symbols and drop NAs
car$country <- iconv(car$country, from="UTF-8", to="LATIN1")
car <- drop_na(car)
```

```{r}
# Column renaming
car <- rename(car, 'customer_name' = 'customer.name')
car <- rename(car, 'customer_e-mail' = 'customer.e.mail')
car <- rename(car, 'annual_salary' = 'annual.Salary')
car <- rename(car, 'credit_card_debt' = 'credit.card.debt')
car <- rename(car, 'net_worth' = 'net.worth')
car <- rename(car, 'car_purchase_amount' = 'car.purchase.amount')
```

```{r}
# create new age group variable
car <- car %>%
                       mutate(age_group =
                                case_when(
                                  age >= 20 & age <= 25 ~ '20 - 25',
                                  age >= 26 & age <= 30 ~ '26 - 30',
                                  age >= 31 & age <= 35 ~ '31 - 35',
                                  age >= 36 & age <= 40 ~ '36 - 40',
                                  age >= 41 & age <= 45 ~ '41 - 45',
                                  age >= 46 & age <= 50 ~ '46 - 50',
                                  age >= 51 & age <= 55 ~ '51 - 55',
                                  age >= 56 & age <= 60 ~ '56 - 60',
                                  age >= 61 & age <= 65 ~ '61 - 65',
                                  TRUE ~ '66 +'
                                ))
```

```{r}
# create male and female variable
car <- car %>% mutate(gender_mf = case_when(
  gender == 0 ~ 'Female',
  gender == 1 ~ 'Male'
))
```

```{r}
# create continent/region variable
library(countrycode)
car$continent <- countrycode(sourcevar = car[, "country"],
                            origin = "country.name",
                            destination = "region")

```



***Data Exploration/Visualization***

# Data View
```{r}
# glimse of first 6 rows
head(car)
```
```{r}
# show all column names
names(car)
```
```{r}
# rows and columns
dim(car)
```
```{r}
# structure
str(car)
```
```{r}
# summary statistics
summary(car)
```


# Summary of individual variables
```{r}
# country
car %>% group_by(country) %>%
  summarise(n_customers = n()) %>%
  arrange(desc(n_customers))
```
Bolivia, Isreal, and Mauritania had the most amount of customers at 6. 


```{r}
# gender
car %>% tabyl(gender)
```
Female had a count of 247 (49.4%) and Male had a count of 253 (50.06%)

```{r}
# age
car %>% group_by(age_group) %>%
  summarise(n_customers = n(),
            average_value = mean(age),
            max_value = max(age),
            min_value = min(age)) %>%
  arrange(desc(n_customers))
```
The age group that has the most customers is 66 + at 104, 41-45 in second at 97. the age group that has the lowest customers is 20-25 at 2.

```{r}
# annual salary
summary(car$annual_salary)
```
The minimum annual salary is 20,000 and the max is 100,000

```{r}
# credit card debt
summary(car$credit_card_debt)
```
The minimum credit card debt is 100 and the max is 20,000

```{r}
# car purchase amount
summary(car$car_purchase_amount)
```
The minimum car purchase amount is 9,000 and the max is 80,000


# Visualizations


Overall correlation of all variables
```{r, fig.width=9, fig.height= 8}
car_numeric <- select(car, c('gender','age', 'annual_salary', 'credit_card_debt', 'net_worth', 'car_purchase_amount'))

#round all variables to whole numbers
car_numeric$age <- round(car_numeric$age,0)
car_numeric$annual_salary <- round(car_numeric$annual_salary,0)
car_numeric$credit_card_debt <- round(car_numeric$credit_card_debt,0)
car_numeric$net_worth <- round(car_numeric$net_worth,0)
car_numeric$car_purchase_amount <- round(car_numeric$car_purchase_amount,0)

corrplot(cor(car_numeric),type = 'upper', method = 'color', diag = FALSE,
         tl.srt = 45, addCoef.col = 'black',
         main = 'Correlation Between Numerical Variables',  mar=c(0,0,1,0))
```
The variables that have the highest correlation:
age & car purchase amount: 64%
annual salary & car purchase amount: 62%
net worth & car purchase amount: 49%


Is there a linear relationship between age and car purchase amount ? How does gender affect this?
```{r, fig.height=8, fig.width=12}
# scatter plot
ggplot(data = car, mapping = aes(x = age,  y = car_purchase_amount, color = gender_mf)) +
       geom_point() +
  geom_smooth(color = 'red') +
  scale_y_continuous(labels = comma) +
  scale_color_manual(values = c('darkorchid2','chartreuse3')) + 
         labs(title = "Correlation Between Age & Gender to Car Purchase Amount",
              subtitle = "Car Purchase Amount is in USD",
            x = "Customer Age",
            y = "Car Purchase Amount",
            color = 'Gender')
```
There seems to be a positive linear relationship between customer age and car purchasing amount. As ago goes up, the amount at which the car purchased goes up as well. There
does not seem to be many outliers either. The relationship is strong. Gender does not seem to change the relationship very much. It seems that there is an even mix of male and female at all values for both axis.


What is the relationship between salary, car purchase amount by age group?
```{r, fig.height=8, fig.width=12}
ggplot(data = car, mapping = aes(x = annual_salary,
                                    y = car_purchase_amount, fill = age_group)) +
  geom_boxplot() +
  scale_x_continuous(labels=comma) +
  scale_y_continuous(labels=comma) +
           labs(title = "Box plot of Salary and Car Purchase Amount by Age Group",
                subtitle = "Values are in USD",
            x = "Customer Annual Salary",
            y = "Customer Car Purchase Amount",
            fill = "Age Group")
```
It seems that the age group of 46-50 has the highest annual salary. This makes sense,
this is usually when a person is at the height of thier career, so thier salary would be at the highest.61-65 has the highest car purchase value.
This makes sense as well, as a person get's older thier savings usually grows and they are able to pay off other purchases. Marketers may try to 
market higher end value cars to these demogrpahics. The lowest car purchase amount is 20-25. Which makes sense, these are people right out of highschool and college
so thier savings are jobs may not be that good.


Counts of each age group's purchase of cars
```{r, fig.height=8, fig.width=12}
ggplot(data = car, aes(x = age_group, fill = gender_mf)) +
    geom_bar(stat = "count") +
    labs(title = "Amount of Cars Purchased by Age Group & Gender",
         x = "Age Group", y = "Cars Purchased",
         fill = "Gender")
```
This plots give more insight into the amount of car purchased by each age group. 66+ had the most
around 105 and then 41-45 at around 95.


Median income and car purchase price by continent
```{r, fig.height=8, fig.width=12}
library(scales)
ggplot(data = car,
       mapping = aes(x = annual_salary, y = car_purchase_amount, fill = continent)) +
  geom_point() +
  scale_x_continuous(labels=comma) +
  scale_y_continuous(labels=comma) +
  geom_smooth(method = "loess", color = 'red') +
  labs(title = "Association of Annual Salary and Car Purchase Amount by Region",
       subtitle = "Values are in USD",
       x = "Customer Annual Salary", 
       y = "Customer Car Purchase Amount",
       fill = "Continent") +
  theme_bw()
```
This plot gives more insight into how annual salary and car purchase amount are associated by region. It looks like Latin America & Carribean
have the lowest annual salary around 20,000 with it maxing around 90,000. Europe and Center Asia has the highest salary at 100,000. North America
has the highest car purchase amount around 90,000.


Is there a linear relationship between credit card debt and car purchase amount ? How does gender affect this?
```{r, fig.height=8, fig.width=12}
# scatter plot
ggplot(data = car, mapping = aes(x = credit_card_debt,  y = car_purchase_amount, color = gender_mf)) +
       geom_point() +
  geom_smooth(color = 'red') +
  scale_x_continuous(labels=comma) +
scale_y_continuous(labels=comma) +
  scale_color_manual(values = c('deeppink','blue3')) + 
         labs(title = "Correlation Between Credit Card Debt & Gender to Car Purchase Amount",
              subtitle = "Values are in USD",
            x = "Customer Credit Card Debt",
            y = "Customer Car Purchase Amount",
            color = 'Gender')
```
Thre does not seem to be much linear association between Car purchase amount and credit card debt. Those with low credit card debt and those with high
credit card debt seem to purhase similar values of cars. Most of the customers have around 5,000 - 15,000 of credit card debt this does not seem to change the value
of car purchased, which is around 30,000 - 50,000

Is there a linear relationship between car purchase amount and net worth?
```{r, fig.height=8, fig.width=12}
ggplot(car, aes(net_worth, car_purchase_amount))+
  geom_jitter(width = .5, size=1, color = 'blue') +
  geom_smooth(color = 'red') +
  scale_x_continuous(labels=comma) +
scale_y_continuous(labels=comma) +
  labs(y="Customer Car Purchase Amount", 
       x="Customer Networth", 
       title="Correlation Between Customer Networth and Car Purchase Amount",
       subtitle = "Values are in USD")
```
It seems that as networth goes up so does car purchase amount. This makes sense those with a higher networth would usually have more disposable cash. This 
would allow them purchase higher valued cars. Those with a higher networth may also have a status or image to keep up with. They are known to have a high 
net worth, so they must present them selfs in nicer higher valued vehicles. 



***Model Creation/Pre-processing***


# Multi-Linear Regression all variables
We will be using gender, age, salary, credit card debt and wet worth to predict car purchase amount

# Split
```{r}
library(rsample)
# Specify the set.seed option to reproduce the random sampling
set.seed(100)

# Generate the rsplit object
car_split <- initial_split(car_numeric, prop = 0.60)

# Extract the training dataset from split object
car_training <- car_split %>% 
  training()
# Extract the validation dataset from split object
car_valid <- car_split %>% 
  testing()
```

```{r}
# run linear regression model
car_reg <- lm(car_purchase_amount~., data=car_training,na.action=na.exclude)
summary(car_reg)
```


By looking at the regression output we can see which variables have the most
affect on car purchasing price, we are able to see this through the t and p values. 
The variable that has the highest correlation is age the next
variable is annual salary, then net_worth . These values all make sense 
from a general stand poiint. Those who have a higher annual salary, are most likely older and 
will also have a higher net worth. With these combination of variables they more likely to be able
to afford and purchase a nicer higher priced vehicle. This is why the t-values are all positive. The 
same can be said for lower priced vehicles. Someone with a lower annual salary, lower age and lower net
worth is likely to buy a lower priced vehicle. Gender and credit card debt do not seem to have
much correlation on car purchasing price. With an R sqaured values of 0.9996 , we can
make the assumption that most of the variation in the data is explained from these varibles alone.
meaning that this model holds true and serves a good representation of how the data acts in a real enviroment. This also shows the model has very high accuracy.

```{r}
pred_t <- predict(car_reg)
pred_v <- predict(car_reg, newdata=car_valid)
```
```{r}
## evaluate performance
# training
accuracy(pred_t, car_training$car_purchase_amount)
# validation
accuracy(pred_v, car_valid$car_purchase_amount)
```



# Predicting new values

Here we will read in the first 5 customers. The point of this is to see if our model
will predict the same or similar car purchase amount for these customers. We can compare the model output
to the actual.
```{r}
first_five_customer <- data.frame(gender=c(0,0,1,1,1), 
                           age=c(42,41,43,58,57), 
                           annual_salary=c(62812 ,66647, 53799, 79370, 59729), 
                           credit_card_debt=c(11609, 9573, 11160, 14426, 5359),
                           net_worth=c(238961, 530974, 638467, 548599, 560304))
predict(car_reg, newdata=first_five_customer)
```

Actual 1: 35321 Predicted 1: 35437 Difference: 116
Actual 2: 45116 Predicted 2: 45224 Difference: 108
Actual 3: 42926 Predicted 3: 42823 Difference: 103
Actual 4: 67422 Predicted 4: 67238 Difference: 184
Actual 5: 55915 Predicted 5: 55616 Difference: 299

It seems the model was able to predicted pretty closely to the original amount.
This shows the model is working correctly and pretty accurately.

we can now create 5 new customers and see what thier car purchase amount would be. Here is 
where the model can be used in a real world application with new data. Someone interested like
a marketing specialist or care sales person can input certain demographics or customers
and change thier tactics based on the output.

```{r}
new_customer <- data.frame(gender=c(0,1,0,0,1), 
                           age=c(34,88,55,39,22), 
                           annual_salary=c(56784 ,44444, 20000, 110000, 96555), 
                           credit_card_debt=c(11655, 95473, 160, 2226, 6659),
                           net_worth=c(200000, 546787, 864233, 10000, 1005450))
predict(car_reg, newdata=new_customer)
```

***Model Results/Visualization***

```{r, fig.height= 10, fig.width=12}
check_model(toyota_reg)
```
```{r}
library('sjPlot')
plot_model(car_reg, type = 'diag', show.values = TRUE)
```



***Analysis/Findings***
Looking at the plot_model output gives us a lot of information about how the variables are related.
The first plot shows variance inflation factors. Which shows if the model is affected negatively by
multicolinearity. Are values are under 1 for all variables which shows there is not redundant
information coming from these variables when used with each other. The non-nomality residual plot
shows the predictied values are closely related to the observed values. Meaning the regression
model performed well in comparission to the actual data. The model only has a few outliers
and follows a normal distribution on the next plot. Overall, this model performs pretty good.
This model could be used to target specific groups with certain vehicles. This could save 
lots of time, money and man power. If a customer or demographic is only likely to 
purchase a vehicle less than 30,000 there is no need to try and get them to purchase a vehicle higher
than that. 

