---
title: "739 Project"
author: "Hugh O'Hara, Megan Parmer, Devin Rohler"
date: "11/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load libraries
library(tidyverse)
library(janitor)
library(lubridate)
library(GGally)
library(psych)
library(corrplot)
```

```{r}
# import
setwd("/Users/Devin/Desktop/School Work/Graduate/Fall 2022/Advanced Datamining")
car= read.csv("car_purchasing.csv", encoding = "UTF-8")
```

***Project Explanation***
This project will be using a dataset with information on vehicle sales from a wide variety of people and attirbutes.  This data set can bring in lots of information about marketing, sales and much more. A lot of insight can be gained for car salespersons, car manufacturers, loan agencies and marketing specialists.
We will be performing machine learning models to see how accuretely we can predict car purchase price based on a customer's attributes. In terms of car salespersons, they could use this model to target certain customers.  Marketing and sales tactics can benefit very much from being able to know exactly what type of customer to go after for certain valued cars. This will save time, effort and money. There is a lot more insight that can be gained once worked through exploration and modeling. 


***Data Cleansing***

```{r}
# clean symbols and drop NAs
car$country <- iconv(car$country, from="UTF-8", to="LATIN1")
car <- drop_na(car)
```

```{r}
# Column renaming
car <- rename(car, 'customer_name' = 'customer.name')
car <- rename(car, 'customer_e-mail' = 'customer.e.mail')
car <- rename(car, 'annual_salary' = 'annual.Salary')
car <- rename(car, 'credit_card_debt' = 'credit.card.debt')
car <- rename(car, 'net_worth' = 'net.worth')
car <- rename(car, 'car_purchase_amount' = 'car.purchase.amount')
```

```{r}
# create new age group variable
car <- car %>%
                       mutate(age_group =
                                case_when(
                                  age >= 20 & age <= 25 ~ '20 - 25',
                                  age >= 26 & age <= 30 ~ '26 - 30',
                                  age >= 31 & age <= 35 ~ '31 - 35',
                                  age >= 36 & age <= 40 ~ '36 - 40',
                                  age >= 41 & age <= 45 ~ '41 - 45',
                                  age >= 46 & age <= 50 ~ '46 - 50',
                                  age >= 51 & age <= 55 ~ '51 - 55',
                                  age >= 56 & age <= 60 ~ '56 - 60',
                                  age >= 61 & age <= 65 ~ '61 - 65',
                                  TRUE ~ '66 +'
                                ))
```

```{r}
# create male and female variable
car <- car %>% mutate(gender_mf = case_when(
  gender == 0 ~ 'Female',
  gender == 1 ~ 'Male'
))
```

```{r}
# create continent variable
library(countrycode)
car$continent <- countrycode(sourcevar = car[, "country"],
                            origin = "country.name",
                            destination = "region")

```


***Data Exploration/Visualization***

# Data View
```{r}
# glimse of first 6 rows
head(car)
```
```{r}
# show all column names
names(car)
```
```{r}
# rows and columns
dim(car)
```
```{r}
# structure
str(car)
```
```{r}
# summary statistics
summary(car)
```


# Summary of individual variables
```{r}
# country
car %>% group_by(country) %>%
  summarise(n_customers = n()) %>%
  arrange(desc(n_customers))
```
Bolivia, Isreal, and Mauritania had the most amount of customers at 6. 


```{r}
# gender
car %>% tabyl(gender)
```
Female had a count of 247 (49.4%) and Male had a count of 253 (50.06%)

```{r}
# age
car %>% group_by(age_group) %>%
  summarise(n_customers = n(),
            average_value = mean(age),
            max_value = max(age),
            min_value = min(age)) %>%
  arrange(desc(n_customers))
```
The age group that has the most customers is 66 + at 104, 41-45 in second at 97. the age group that has the lowest customers is 20-25 at 2.

```{r}
# annual salary
summary(car$annual_salary)
```
The minimum annual salary is 20,000 and the max is 100,000

```{r}
# credit card debt
summary(car$credit_card_debt)
```
The minimum credit card debt is 100 and the max is 20,000

```{r}
# car purchase amount
summary(car$car_purchase_amount)
```
The minimum car purchase amount is 9,000 and the max is 80,000


# Visualizations


Overall correlation of all variables
```{r, fig.width=9, fig.height= 8}
car_numeric <- select(car, c('gender','age', 'annual_salary', 'credit_card_debt', 'net_worth', 'car_purchase_amount'))
corrplot(cor(car_numeric),type = 'upper', method = 'color', diag = FALSE,
         tl.srt = 45, addCoef.col = 'black',
         main = 'Correlation Between Numerical Variables',  mar=c(0,0,1,0))
```
The variables that have the highest correlation:
age & car purchase amount: 64%
annual salary & car purchase amount: 62%
net worth & car purchase amount: 49%


Is there a linear relationship between age and car purchase amount ? How does gender affect this?
```{r}
# scatter plot
ggplot(data = car, mapping = aes(x = age,  y = car_purchase_amount, color = gender_mf)) +
       geom_point() +
  geom_smooth(color = 'red') +
  scale_color_manual(values = c('darkorchid2','chartreuse3')) + 
         labs(title = "Correlation Between Age & Gender to Car Purchase Amount",
            x = "Customer Age",
            y = "Car Purchase Amount",
            color = 'Gender')
```
There seems to be a positive linear relationship between customer age and car purchasing amount. As ago goes up, the amount at which the car purchased goes up as well. There
does not seem to be many outliers either. The relationship is strong. Gender does not seem to change the relationship very much. It seems that there is an even mix of male and female at all values for both axis.


What is the relationship between salary, car purchase amount by age group?
```{r}
ggplot(data = car, mapping = aes(x = annual_salary,
                                    y = car_purchase_amount, fill = age_group)) +
  geom_boxplot() +
           labs(title = "Box plot of Salary and Car Purchase Amount by Age Group",
            x = "Customer Annual Salary",
            y = "Customer Car Purchase Amount",
            fill = "Age Group")
```
It seems that the age group of 46-50 has the highest annual salary. This makes sense,
this is usually when a person is at the height of thier career, so thier salary would be at the highest.61-65 has the highest car purchase value.
This makes sense as well, as a person get's older thier savings usually grows and they are able to pay off other purchases. Marketers may try to 
market higher end value cars to these demogrpahics. The lowest car purchase amount is 20-25. Which makes sense, these are people right out of highschool and college
so thier savings are jobs may not be that good.


Counts of each age group's purchase of cars
```{r}
ggplot(data = car, aes(x = age_group, fill = gender_mf)) +
    geom_bar(stat = "count") +
    labs(title = "Amount of Cars Purchased by Age Group & Gender",
         x = "Age Group", y = "Cars Purchased",
         fill = "Gender")
```
This plots give more insight into the amount of car purchased by each age group. 66+ had the most
around 105 and then 41-45 at around 95.


Median income and car purchase price by continent

```{r, fig.height=8, fig.width=12}
ggplot(data = car,
       mapping = aes(x = annual_salary, y = car_purchase_amount, fill = continent)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "Association of Annual Salary and Car Purchase Amount by Region", 
       x = "Customer Annual Salary", 
       y = "Customer Car Purchase Amount",
       fill = "Continent") +
  theme_bw() +
  scale_x_continuous(labels = comma)

```
This plot gives more insight into how annual salary and car purchase amount are associated by region. It looks like Latin America & Carribean
have the lowest annual salary around 20,000 with it maxing around 90,000. Europe and Center Asia has the highest salary at 100,000. North America
has the highest car purchase amount around 90,000.

***Model Creation/Pre-processing***






***Model Results***





***Analysis/Findings***





***Future Work***